Supporting Information : Effect of copula on extinction risk

### 2019-03-13 MainText.Rmd
methods : writing age structure model completed

<!--stochastic matrix model : age structured-->
To investigate how copula structure in noisy environment affect the extinction risk of
populations in a single patch age-structured model,
we considered a simple stochastic age-structured model (Eq. \ref{eq_age_str}) 
with two age classes: 
one is egg ($E$) and the other is adult ($A$). We assumed
the population was semelperous, i.e., adult ($A$) went through the
reproductive episode only once before death. Eggs and adults in the 
next generation depend upon the fecundity of adult ($f_a \geq 0$) 
and survival of eggs ($0 \leq s_e \leq 1$) from present generation, respectively.
The environmental stochasticity in the model was embedded in 
the fluctuating vital rates ($f_a$ and $s_e$). We considered the fecundity ($f_a$)
and survival ($s_e$) rates were random draw from a gamma ($\Gamma(k,\theta)$) and
a beta ($\mathcal{B}(\alpha,\beta)$) distribution, respectively. These distribution 
could be generated from iid noises showing specific tail-dependence (left or right tail)
between fecundity and survival rates. Eq. \ref{eq_vitalrates} represent the stochasticity 
in the vital rates. Demographic stochasticity was also taken into consideration when 
we included the density dependence effect in one of the vital rate, namely, fecundity ($f_a$).
If $\Omega = \infty$ in Eq. \ref{eq_vitalrates}, the projection matrix was independent
of populations from any age-class and the model was then called to be "linear", whereas
its "nonlinear" version had finite $\Omega$.

@@

We repeated the above analysis considering the vital rates coming from a 
an asymmetric copula (Appendix \ref{SM-age_str_vitalrates}) with specific 
left (Clayton, Survival Gumbel)
or right (Survival Clayton, Gumbel) tail dependence
or from an extreme tail-depentdent one, when vital rates were considered to be affected
by extreme climatic events.

>>>>

<!--stochastic matrix model : age structured-->
To investigate how copula structure in noisy environment affect the extinction risk of
populations in a single patch age-structured model,
we considered a simple stochastic age-structured model (Eq. \ref{eq_age_str}) 
with two age classes(egg $E$ and adult $A$). We assumed
the population was semelperous, i.e., adult ($A$) went through the
reproductive episode only once before death. Eggs and adults in the 
next generation depend upon the fecundity of adult ($f_a \geq 0$) 
and survival of eggs ($0 \leq s_e \leq 1$) from present generation, respectively.
The environmental stochasticity in the model was embedded in 
the fluctuating vital rates ($f_a$ and $s_e$). We considered the fecundity ($f_a$)
and survival ($s_e$) rates were random draw from a gamma ($\Gamma(k,\theta)$) and
a beta ($\mathcal{B}(\alpha,\beta)$) distribution, respectively. These distribution 
could be generated from iid noises showing specific tail-dependence (left or right tail)
between fecundity and survival rates. Eq. \ref{eq_vitalrates} represent the stochasticity 
in the vital rates. We included the density dependence effect in one of the vital rate,
namely, fecundity ($f_a$). If $\Omega = \infty$ in Eq. \ref{eq_vitalrates}, 
the projection matrix would be independent
of populations from any age-class and the model was then called to be "linear", whereas
its "nonlinear" version had finite $\Omega$.

@@

We repeated the above analysis considering the vital rates coming from a 
an asymmetric copula with specific 
left (Clayton, Survival Gumbel)
or right (Survival Clayton, Gumbel) tail dependence
or from an extreme tail-depentdent one, 
when vital rates were considered to be affected
by extreme climatic events. A $N \times 2$ matrix, M, was generated
using `BiCopSim` function
from `VineCopula` package 
for each of the left 
(Clayton, Survival Gumbel) and right 
(Survival Clayton, Gumbel) tail-dependent bivariate copulas, 
where $N = numsteps(=25) \times numsims(=10000)$.
We kept the same Kendall's correlation ($\tau = 0.6$) for 
each bivariate copula from one parameter family which used its parameter accordingly
(see `iTau` function from `copula` package). Entries within each clumn of
matrix $M$, were thus 
independent of each others but 
the two columns of $M$ had a specified Kendall's 
correlation ($\tau = 0.6$) and underlying tail-dependence. 
Similarly, for extreme-tail dependence
(either left or right tail) a matrix M with two column was generated as described in 
Appendix \ref{SM-extr_tail_dep}.

We made the first column of matrix 
$M$ as gamma ($\Gamma(k,\theta)$) distributed. 
$k = 2$ and $\theta = 1$ were shape and scale parameters, 
respectively. This column represented the distribution of fecundity of the adult population, $f_a$.
For a given $\theta$, increasing 
$k$ indicates distribution will shift towards higher fecundity which might 
reduce the extinction risk if the other parameters
remain unchanged. We took the second column of $M$ and applied a 
beta ($\mathcal{B}(\alpha,\beta)$) transformation. $\alpha =3$ and
$\beta = 2$ were the shape parameters which could control the distribution 
of the survival rates of eggs, $s_e$. 
For each $25$ timesteps from each of $10000$ simulations,
the vital rates (fecundity $f_a$ and survival rates $s_e$) 
of the stochastic projrction matrix in model (\ref{eq_age_str})
were drawn pairwise from each row of matrix $M$.
[`sim_age_str`](https://github.com/sghosh89/ERC/blob/master/ext_risk_cop_age_str.R) 
is the function we used to calculate extinction risk for age structured model.

### 2019-03-13 SuppMat.Rmd
methods : writing age structure model completed

# Vital rates in a stochastic age-structured model having specific tail-dependence \label{age_str_vitalrates}

A $N \times 2$ matrix, M, was generated for each of the left (Clayton, Survival Gumbel) and right 
(Survival Clayton, Gumbel) tail-dependent bivariate copulas, where $N = numsteps(=25) \times numsims(=10000)$.
We kept the same Kendall's correlation ($\tau = 0.6$) for each bivariate copula which used parameter accordingly
(see `iTau` function from `copula` package). For fecundity of adults, we transformed the first column of matrix 
$M$ as gamma ($\Gamma(k,\theta)$) distributed, $k = 2$ and $\theta = 1$ were shape and scale parameters, 
respectively. For a given $\theta$, increasing 
$k$ indicates distribution will shift towards higher fecundity which might 
reduce the extinction risk if the other parameters
remain unchanged. For survival rates of eggs, we took the second column of $M$ and applied a 
beta ($\mathcal{B}(\alpha,\beta)$) transformation, $\alpha =3$ and
$\beta = 2$ were the shape parameters which could control the distribution 
of the survival rates of eggs. Entries within each clumn of matrix $M$, were thus independent of each others but 
the two columns of $M$ had a specified Kendall's correlation ($\tau = 0.6$) and underlying tail-dependence. 
For each $25$ timesteps from each of $10000$ simulations, the vital rates (fecundity and survival rates) 
of the stochastic projrction matrix in model (\ref{MT-eq_age_str}) were drawn pairwise from each row of matrix $M$.

For extreme tail-dependence algorithm,

>>>>

<!--
# Vital rates in a stochastic age-structured model having specific tail-dependence \label{age_str_vitalrates}
A $N \times 2$ matrix, M, was generated for each of the left (Clayton, Survival Gumbel) and right 
(Survival Clayton, Gumbel) tail-dependent bivariate copulas, where $N = numsteps(=25) \times numsims(=10000)$.
We kept the same Kendall's correlation ($\tau = 0.6$) for each bivariate copula which used parameter accordingly
(see `iTau` function from `copula` package).-->

# Generating matrix with extreme tail dependence  \label{extr_tail_dep}

For extreme tail-dependence algorithm, a $N \times n$ matrix, M was generated of independent 
standard-normal random draws (in general), where $N = numsteps(=25) \times numsims(=10000)$.
Then, each row of
M was identified for a first set of modifications
randomly and independently with a 50\% chance; for those rows identified, every
element of the row was replaced with the absolute
value of the first element of the row (including the first element itself).
For other rows, every element in the row was replaced by the negative of 
its own absolute value. This produces right-tail dependent
matrix, normally distributed for each column. If left-tail 
dependent noise is desired, take the negative of the entire matrix.
Our [code](https://github.com/sghosh89/ERC/blob/master/ExtremeTailDep.R) for
generating extreme tail dependence matrix is flexible enough, so that 
matrix M can be generated initially with independent random draws from a normal 
distribution with a specified mean ($m$) and standard deviation $(sd)$.
